---
title: "Model Plan"
author: "Levi Carr"
format: html
html:
   embed-resources: true
---
```{r, include=FALSE}
library(ggplot2)
library(ggdag)
library(dagitty)
```

## Causal Diagram
```{r}
causal_diag <- dagitty('dag {
"Cloud Top Pressure" [exposure,pos="0.592,-0.460"]
"Droplet Nucleus Composition" [latent,pos="-1.468,-1.183"]
"Effective Radius" [outcome,pos="-0.300,-0.082"]
"Ice Fraction" [exposure,pos="-0.813,-1.179"]
"Cloud Top Pressure" -> "Effective Radius" [pos="0.680,-0.496"]
"Cloud Top Pressure" -> "Ice Fraction"
"Droplet Nucleus Composition" -> "Ice Fraction"
"Ice Fraction" -> "Effective Radius" [pos="-0.791,-1.045"]
}'
)

ggdag(causal_diag)
```
## Initial Model & Priors
 $$
 R_{eff} \sim \mathrm{Gamma}(\mu_i, \sigma)
 $$
 $$
 \ln(\mu_i) = \beta_0 + \beta_1 \;\text{ctp_ice.Z} + \beta_2 \;\text{fraction_ice}
 $$
 $$
 \beta_0 \sim \mathrm{Cauchy}(\ln(125), \ln(50))
 $$
 $$
 \beta_1 \sim \mathrm{Cauchy}(0, 1)
 $$
 $$
 \beta_2 \sim \mathrm{Normal}(0,1)
 $$
 $$
 \sigma \sim \mathrm{Lognormal}(0,1)
 $$

From prior information, ice clouds tend to have particles with wide distributions in radius, with an apparent peak near $\mu_{R_{eff}} \approx$ 125 $\mu$m and $\sigma_{R_{eff}} \approx$ 50 $\mu$m (cf. DOI: 10.1002/2017EA000296, pg 11). Likewise, ctp_ice spans 2 decades of pressure, but I do not have a reasonable guess for what the mean and standard deviation should be. Thus, I choose to normalize ctp_ice and use an uninformative prior. Fraction_ice is bounded between 0 and 1, but it too does not seem to have previously known information. Lastly, I chose the Lognormal sd prior as to eliminate negative standard deviations. 

In the next week, I will design a prior predictive code and examine it to see the choice of prior and critique its initial values.